A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-06 2022/OM&S/paper.tex
1. section 3

>>>>

According to ARL applied to solving the problem (\ref{problem_fi}) its first $l \geq 2$  trials are executed at arbitrary points $t^1, \dots, t^l$ of the interval $[\alpha, \beta]$ including points $\alpha$ and $\beta$, and the function values $\nu^1, \dots, \nu^l$ are evaluated at these points, i.e., $\nu^s=\varphi(t^s)$, $1 \leq j \leq l$.

In order to choose a point $t^{n+1}$ of a new $(n+1)$-th trial after completing $n \geq l$ trials it is necessary to implement the following actions.

1. The points $t^1, \dots, t^n$ of completed trials and boundary points of the search interval $[\alpha, \beta]$ are ordered in increasing order and renumbered by subscripts, i.e.,

\begin{equation}\label{eq:15}
t_1 = \alpha < t_1 < \dots < t_{n-1} < t_n = \beta.
\end{equation}

The values $\nu_s = \varphi(t_s)$ are juxtaposed to the points  $t_s$, $1 \leq s \leq n$.

2. A real values $R(q)$ are assigned to all the intervals $(t_{q-1}, t_q)$, $2 \leq q \leq n$, formed by neighboring points from (\ref{eq:15}). $R(q)$ is called \textit{the characteristic} of the interval $(t_{q-1}, t_q)$ and is computed in accordance with the following rule.

If number of completed trials $2 \leq n \leq J$, where $J$ is a predefined number, then 
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise, when $n > J$,
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \sqrt{(\nu_q - \varphi^*(n))(\nu_{q-1} - \varphi^*(n))} +M(1.5)^{-\sigma}\right)^{-1}.
\end{equation}

Here
\begin{equation}\label{eq:18}
\begin{matrix}
B(q) = m (t_s - t_{s-1})^{1/N} + (\nu_q - \nu_{q-1})^2 / (m (t_s - t_{s-1})^{1/N}) -2(\nu_{q-1} + \nu_q), \\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{matrix}
\end{equation}
is minimal computed value of the objective function $\varphi(t)$ at trial points,

\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|\nu_s - \nu_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}

is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}

is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

3. Among all intervals, the interval   with the highest characteristic is selected
\begin{equation}\label{eq:21}
R(p) = \max_{2 \leq s \leq n} {R(s)}
\end{equation}

4. The new $(n+1)$-th trial is performed at the point
\begin{equation}\label{eq:22}
t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(\nu_p-\nu_{p-1}) \left( \frac{|\nu_p-\nu_{p-1}|}{M} \right)^N \frac{1}{2r} 
\end{equation}
the value $\nu^{n+1} = \varphi(t^{n+1})$ is computed and the iteration number $n$ is increased by 1.

The method ALR described above is a modification of the base information-statistical algorithm [Strongin+Sergeyev, Sergeyev+Strongin] for minimization functions satisfying the H{\"o}lder condition. The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).

Like its prototype, ARL can apply as the stopping condition the inequality
\begin{equation}\label{eq:23}
(t_p-t_{p-1})^{1/N} < \varepsilon
\end{equation}
i.e., completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.

The sufficient condition of convergence to global minima in the problem (\ref{problem_fi}) with the objective function under H{\"o}lder property with constant $H$ is
\begin{equation}\label{eq:24}
m > 2^{2 - 1 / N}H.
\end{equation}

If $\varphi(t) = f(x(t)) $ where $x(t)$ is the Peano-type curve and $f(x)$ meets (\ref{lip_1}), then $H=2L\sqrt{N+3}$ and the sufficient condition takes the form 
\begin{equation}\label{eq:25}
m > 2^{3 - 1 / N}L\sqrt{N+3}.
\end{equation}

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
